% A LaTeX template for MSc Thesis submissions to 
% Politecnico di Milano (PoliMi) - School of Industrial and Information Engineering
%
% S. Bonetti, A. Gruttadauria, G. Mescolini, A. Zingaro
% e-mail: template-tesi-ingind@polimi.it
%
% Last Revision: October 2021
%
% Copyright 2021 Politecnico di Milano, Italy. NC-BY

\documentclass{Configuration_Files/PoliMi3i_thesis}

%------------------------------------------------------------------------------
%	REQUIRED PACKAGES AND  CONFIGURATIONS
%------------------------------------------------------------------------------

% CONFIGURATIONS
\usepackage{parskip} % For paragraph layout
\usepackage{setspace} % For using single or double spacing
\usepackage{emptypage} % To insert empty pages
\usepackage{multicol} % To write in multiple columns (executive summary)
\setlength\columnsep{15pt} % Column separation in executive summary
\setlength\parindent{0pt} % Indentation
\raggedbottom  

% PACKAGES FOR TITLES
\usepackage{titlesec}
% \titlespacing{\section}{left spacing}{before spacing}{after spacing}
\titlespacing{\section}{0pt}{3.3ex}{2ex}
\titlespacing{\subsection}{0pt}{3.3ex}{1.65ex}
\titlespacing{\subsubsection}{0pt}{3.3ex}{1ex}
\usepackage{color}

% PACKAGES FOR LANGUAGE AND FONT
\usepackage[english]{babel} % The document is in English  
\usepackage[utf8]{inputenc} % UTF8 encoding
\usepackage[T1]{fontenc} % Font encoding
\usepackage[11pt]{moresize} % Big fonts

% PACKAGES FOR IMAGES
\usepackage{graphicx}
\usepackage{transparent} % Enables transparent images
\usepackage{eso-pic} % For the background picture on the title page
\usepackage{subfig} % Numbered and caption subfigures using \subfloat.
\usepackage{tikz} % A package for high-quality hand-made figures.
\usetikzlibrary{}
\graphicspath{{./Images/}} % Directory of the images
\usepackage{caption} % Coloured captions
\usepackage{xcolor} % Coloured captions
\usepackage{amsthm,thmtools,xcolor} % Coloured "Theorem"
\usepackage{float}

% STANDARD MATH PACKAGES
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{bm}
\usepackage[overload]{empheq} % For braced-style systems of equations.
\usepackage{fix-cm} % To override original LaTeX restrictions on sizes

% PACKAGES FOR TABLES
\usepackage{tabularx}
\usepackage{longtable} % Tables that can span several pages
\usepackage{colortbl}

% PACKAGES FOR ALGORITHMS (PSEUDO-CODE)
\usepackage{algorithm}
\usepackage{algorithmic}

% PACKAGES FOR REFERENCES & BIBLIOGRAPHY
\usepackage[colorlinks=true,linkcolor=black,anchorcolor=black,citecolor=black,filecolor=black,menucolor=black,runcolor=black,urlcolor=black]{hyperref} % Adds clickable links at references
\usepackage{cleveref}
\usepackage[square, numbers, sort&compress]{natbib} % Square brackets, citing references with numbers, citations sorted by appearance in the text and compressed
\bibliographystyle{abbrvnat} % You may use a different style adapted to your field

% OTHER PACKAGES
\usepackage{pdfpages} % To include a pdf file
\usepackage{afterpage}
\usepackage{lipsum} % DUMMY PACKAGE
\usepackage{fancyhdr} % For the headers
\fancyhf{}

% Input of configuration file. Do not change config.tex file unless you really know what you are doing. 
\input{Configuration_Files/config}

%----------------------------------------------------------------------------
%	NEW COMMANDS DEFINED
%----------------------------------------------------------------------------


%----------------------------------------------------------------------------
%	ADD YOUR PACKAGES (be careful of package interaction)
%----------------------------------------------------------------------------

\usepackage[innercaption]{sidecap}
\usepackage{wrapfig}
\usepackage{csquotes}
\usepackage{listings}

%----------------------------------------------------------------------------
%	ADD YOUR DEFINITIONS AND COMMANDS (be careful of existing commands)
%----------------------------------------------------------------------------

\input{Configuration_Files/custom_commands}

%----------------------------------------------------------------------------
%	BEGIN OF YOUR DOCUMENT
%----------------------------------------------------------------------------

\begin{document}

\fancypagestyle{plain}{%
\fancyhf{} % Clear all header and footer fields
\fancyhead[RO,RE]{\thepage} %RO=right odd, RE=right even
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}}

%----------------------------------------------------------------------------
%	TITLE PAGE
%----------------------------------------------------------------------------

\pagestyle{empty} % No page numbers
\frontmatter % Use roman page numbering style (i, ii, iii, iv...) for the preamble pages

\puttitle{
	title=Programming Environment for \\Human-Drone Interaction: EasyFly, % Title of the thesis
	name=Matteo Plona, % Author Name and Surname
	course=Computer Science Engineering, % Study Programme (in Italian)
	ID  = 952967,  % Student ID number (numero di matricola)
	advisor= Prof. Luca Mottola, % Supervisor name
	%coadvisor={Name Surname}, % Co-Supervisor name, remove this line if there is none
	academicyear={2022-23},  % Academic Year
} % These info will be put into your Title page 

%----------------------------------------------------------------------------
%	PREAMBLE PAGES: ABSTRACT (inglese e italiano), EXECUTIVE SUMMARY
%----------------------------------------------------------------------------
\startpreamble
\setcounter{page}{1} % Set page counter to 1

% ABSTRACT IN ENGLISH
\chapter*{Abstract} 
In recent years, the proliferation of unmanned aerial vehicles, commonly known as drones, has profoundly impacted society.
% This new technology has revolutionized various industries and opened up new possibilities for applications ranging from aerial photography and package delivery to search and rescue operations. 
As drones become increasingly integrated into our daily lives, exploring and understanding the dynamics of their interaction with humans is crucial.

The Human-Drone Interaction (HDI) research field focuses on understanding, designing, and evaluating drone systems for use by or with human users. 
The outputs of this research guide the continuous expansion of this technology.

Programming HDI is usually the field's most complex and expensive task. 
Usually, researchers in this field are unfamiliar with programming at low-level embedded systems like drones.

EasyFly addresses these challenges by offering an accessible programming environment tailored for HDI research. 
It empowers users with varying levels of expertise to experiment with drones.

Its modular approach in hardware and software enhances adaptability and flexibility.
To implement EasyFly, we have targeted a specific typology of unmanned aerial vehicles: nano-drones.
Given their size, this kind of drone facilitates safe and effective human-drone interactions.

The thesis evaluates EasyFly within the Digital Futures Drone Arena project~\cite{dronearena}, where it played a crucial role in an aerial drone testbed and participated in a challenge comparing its efficacy against a lower-level programming alternative.
In the evaluation, we demonstrated the effectiveness of EasyFly in building simple and clean programs for drone applications.

In conclusion, this research contributes to the evolving landscape of HDI, emphasizing accessibility and adaptability in drone applications.

\textbf{Keywords:} EasyFly, Human-drone interaction, programming environment for drones. % Keywords

% ABSTRACT IN ITALIAN
\chapter*{Abstract in lingua italiana}
Negli ultimi anni, la proliferazione di droni, ha avuto un impatto profondo sulla società. 
Questa nuova tecnologia ha rivoluzionato varie industrie e aperto nuove possibilità di applicazioni che vanno dalla fotografia aerea alla consegna di pacchi fino alle operazioni di ricerca e soccorso. 
Con l'integrazione sempre maggiore dei droni nella vita quotidiana, è cruciale esplorare e comprendere la dinamica della loro interazione con gli esseri umani.

Il campo di ricerca dell'Interazione Uomo-Drone (HDI) si concentra sulla comprensione, progettazione e valutazione dei sistemi atti per l'uso di droni da parte degli utenti. 
I risultati di questa ricerca guidano la continua espansione di questa tecnologia.

La programmazione di interazioni uomo-drone è generalmente il processo più complesso ed costoso del settore. 
Di solito, i ricercatori in questo campo non sono familiari con la programmazione a basso livello di embedded systems come i droni.

EasyFly affronta queste sfide offrendo un ambiente di programmazione accessibile su misura per la ricerca nel campo di HDI. 
Consente agli utenti con diversi livelli di competenza di sperimentare con i droni.

Il suo approccio modulare nell'hardware e nel software migliora l'adattabilità e la flessibilità. 
Per implementare EasyFly, ci siamo concentrati su una specifica tipologia di droni: i nano-droni. 
Date le loro dimensioni, questo tipo di drone facilita interazioni sicure ed efficaci tra umani e droni.

La tesi valuta EasyFly all'interno del progetto Digital Futures Drone Arena~\cite{dronearena}, dove ha svolto un ruolo cruciale in una piattaforma di test per droni e ha partecipato a una sfida confrontando la sua efficacia con un'alternativa di programmazione a basso livello. 
Nella valutazione, abbiamo dimostrato l'efficacia di EasyFly nella creazione di programmi semplici e puliti per applicazioni drone.

In conclusione, questa ricerca contribuisce al panorama in evoluzione di HDI, enfatizzando l'accessibilità e l'adattabilità nelle applicazioni drone.

\textbf{Parole chiave:} EasyFly, Interazioni uomo-drone, ambiente di programmazione per droni. % Keywords (italian)

%----------------------------------------------------------------------------
%	LIST OF CONTENTS/FIGURES/TABLES/SYMBOLS
%----------------------------------------------------------------------------

% TABLE OF CONTENTS
\thispagestyle{empty}
\tableofcontents % Table of contents 
\thispagestyle{empty}
\cleardoublepage

%-------------------------------------------------------------------------
%	THESIS MAIN TEXT
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\mainmatter % Begin numeric (1,2,3...) page numbering

% --------------------------------------------------------------------------
% CHAPTERS 
% --------------------------------------------------------------------------
\chapter{Introduction}
\label{ch:intro}
\input{chapters/1_intro/introduction}

\chapter{State of The Art}
\label{ch:soa}
\input{chapters/2_soa/state-of-the-atr.tex}

\chapter{Tools}
\label{ch:tools}
\input{chapters/3_tools/tools.tex}

\chapter{Extended Crazyflie}
\label{ch:ecf}
\input{chapters/4_ecf/extended-crazyflie.tex}

\chapter{Communication Framework}
\label{ch:communication}
\input{chapters/5_communication/communication-framework.tex}

\chapter{Coordination Framework}
\label{ch:coordination}
\input{chapters/6_coordination/coordination-framework.tex}

\chapter{ECF Modules}
\label{ch:modules}
\input{chapters/7_modules/ecf-modules.tex}

\chapter{Conclusions and future developments}
\label{ch:conclusions}
\input{chapters/8_evaluation/evaluation.tex}

%-------------------------------------------------------------------------
%	BIBLIOGRAPHY
%-------------------------------------------------------------------------

\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics
\bibliography{Thesis_bibliography} % The references information are stored in the file named "Thesis_bibliography.bib"

%-------------------------------------------------------------------------
%	APPENDICES
%-------------------------------------------------------------------------

\cleardoublepage
\addtocontents{toc}{\vspace{2em}} % Add a gap in the Contents, for aesthetics

% LIST OF FIGURES
\listoffigures

% LIST OF TABLES
\listoftables

% LIST OF LISTING
\lstlistoflistings

% ACKNOWLEDGEMENTS
\chapter*{Acknowledgements}
Questa tesi è dedicata alle quattro persone più importanti della mia vita: mio figlio, Liam, la mia compagna di vita Angela e i miei genitori Anna e Marco.

A mio figlio, che mi ha dato gioia e sorrisi ogni giorno. 
Sei stato la luce che mi ha guidato in questo lungo percorso. 
Mi hai permesso di vedere il mondo con altri occhi e riempito la mia vita di soddisfazioni.

Alla mia amata Angela, che è stata la mia roccia ed il mio amore.
Il tuo infinito supporto ed incoraggiamento sono stati il fondamento del mio successo sia in questo lavoro che nella vita di tutti i giorni.
Ti sarò per sempre grato per la tua capacità nel rendermi sempre una persona migliore.

Ai miei genitori, Anna e Marco, che hanno indotto in me una passione per lo studio e per la ricerca di nuove conoscenze.
I vostri enormi sacrifici, e la vostra guida mu hanno permesso di raggiungere tutti i miei obiettivi.
In particolare volevo ringraziare te, Mamma, per non aver mai smesso di spronarmi durante il mio percorso fin da quando ero bambino. 
Vi sarò per sempre grato per tutte le opportunità che mi avete donato.

In ultimo vorrei dedicare questa tesi ai miei Nonni, che purtroppo oggi non possono celebrare con me questo fantastico traguardo, pur sapendo che mi guidano ogni giorno grazie ai loro insegnamenti.

\cleardoublepage

\end{document}
